---
title: "Analysis of LCI Thermal Mark Hatchery Otoliths, 2018-2022"
author: "Jonah Bacon"
date: "`r Sys.Date()`"
output: 
  html_document:
      toc: true
      toc_float:
        collapsed: false
        smooth_scroll: true
---

**Revision 2**

**This code treats stratum as inclusive of StatWk. All analysis includes StatWk as a separate dimension of 'strata', as opposed to pooling StatWk's and going to the higher level 'Year'**

This code analyzes data collected from 2018-2022 as part of the Thermal Mark Hatchery Salmon Otoliths in Lower Cook Inlet project (*REGIONAL OPERATIONAL PLAN NO. ROP.CF.2A.2023.XX*). Data, scripts, and associated files are located in the Homer LAN network: ```O:\DCF\SALMON\OTOLITH_STUDIES\1_ANALYSES\2018-2022```. Data are first cleaned, edited (when commercial harvest was misreported or errors in recording occurred), and then analyzed for Pink Salmon followed by identical analysis for Sockeye Salmon.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = "C:/Users/jabacon/Documents/2018-2022")

# Load libraries
library(tidyverse)

```

# Prepare Data
## Data

Load data files from ```O:\DCF\SALMON\OTOLITH_STUDIES\1_ANALYSES\2018-2022\input```.

```{r loaddata}
pink.marks <- read.csv("input/PinkMarks.csv")
reds.marks <- read.csv("input/RedMarks.csv")
harvest <- read.csv("input/tblHarvest.csv")
sample <- read.csv("input/tblHomerSample.csv")
ChangeStatArea <- read.csv("input/ChangeStatArea.csv")
MissingHarvest <- read.csv("input/MissingHarvest.csv")
```

## Clean Data

```{r cleanotolithdata}

pink.marks$StatArea <- as.character(pink.marks$StatArea)
pink.clean <- pink.marks %>% 
  filter(Source %in% c("COMMERCIAL COM PROP","COST RECOVERY")) %>%
  select(-c(Agency,FisheryName,Species,SubDistrict,AStreamCode,Stock,Harvest,SampleDate,SurveySite)) %>% 
  separate(StatArea, into = c("StatArea","Extra"), sep = 5) %>% 
  mutate(StatArea = as.integer(StatArea)) %>% 
  filter(StatArea > 10000) %>% 
  select(-Extra) %>% 
  mutate(# StatArea = if_else(Gear == "SET GILLNET", "24100", StatArea),
         StatArea = ifelse(HomerSampleID == "18PSD45P",24107,StatArea), # Specific correction
         StatArea = ifelse(HomerSampleID == "19GSD086P",24106,StatArea), # Specific correction, don't need any longer?
         Gear = if_else(Gear == "PURSE SEINE", "PS", Gear),
         Gear = if_else(Gear == "SET GILLNET", "SGN", Gear),
         Source = if_else(Source == "COMMERCIAL COM PROP", "CCP", Source),
         Source = if_else(Source == "COST RECOVERY", "HCR", Source)) %>% 
  unite("Year_StatWk_Gear", c(Year,StatWeek,Gear), sep = "_") %>% 
  mutate(Year_StatWk_Gear = as.factor(Year_StatWk_Gear),StatArea = as.factor(StatArea),Source = as.factor(Source),
         Species = "Pink")

reds.clean <- reds.marks %>% 
  filter(Source %in% c("COMMERCIAL COM PROP","COST RECOVERY")) %>%
  select(-c(Agency,FisheryName,Species,SubDistrict,AStreamCode,Stock,Harvest,SampleDate,SurveySite)) %>% 
  separate(StatArea, into = c("StatArea","Extra"), sep = 5) %>% 
  mutate(StatArea = as.integer(StatArea)) %>% 
  filter(StatArea > 10000) %>% 
  select(-Extra) %>% 
  mutate(# StatArea = if_else(Gear == "SET GILLNET", "24100", StatArea),
         Gear = if_else(Gear == "PURSE SEINE", "PS", Gear),
         Gear = if_else(Gear == "SET GILLNET", "SGN", Gear),
         Source = if_else(Source == "COMMERCIAL COM PROP", "CCP", Source),
         Source = if_else(Source == "COST RECOVERY", "HCR", Source),
         StatArea = ifelse(HomerSampleID == "18PSD06S",24190,StatArea),
         StatArea = ifelse(HomerSampleID == "18PSD03S",24190,StatArea),
         StatArea = ifelse(HomerSampleID == "18GHC10S",24115,StatArea),
         StatArea = ifelse(HomerSampleID == "18GPG11S",24130,StatArea),
         StatArea = ifelse(HomerSampleID == "18GPG06S",24130,StatArea),
         StatArea = ifelse(HomerSampleID == "18GPG03S",24130,StatArea),
         StatArea = ifelse(HomerSampleID == "18PSD06S",24190,StatArea),
         StatArea = ifelse(HomerSampleID == "18PSD03S",24190,StatArea)) %>% 
  unite("Year_StatWk_Gear", c(Year,StatWeek,Gear), sep = "_") %>% 
  mutate(Year_StatWk_Gear = ifelse(HomerSampleID == "22GMX005S","2022_25_SGN",Year_StatWk_Gear),
         Year_StatWk_Gear = as.factor(Year_StatWk_Gear),StatArea = as.factor(StatArea),Source = as.factor(Source),
         Species = "Sockeye")

otolith.TM.data <- pink.clean %>% 
  select(HomerSampleID,Year_StatWk_Gear,Species,Source,StatArea,Preps,NotMarked,Marked,LCI,PWS,KOD,Other) %>% 
  full_join(select(reds.clean, c(HomerSampleID,Year_StatWk_Gear,Species,Source,StatArea,Preps,NotMarked,Marked,LCI,PWS,KOD,Other)))
```

```{r cleanharvestdata}
harvest.df <- harvest %>% 
  select(-c(ID,DOL.Month,District,Subdistrict,Stat.Area.Name,Landed.Weight..sum.,Business.Name)) %>% 
  filter(Species.Name %in% c("salmon, sockeye","salmon, pink"),Gear.Name %in% c("Purse seine","Set gillnet"),Harvest.Name %in% c("State managed fishery","Private Hatchery - Fishing")) %>% 
  mutate(Gear.Name = if_else(Gear.Name == "Purse seine", "PS", Gear.Name),
         Gear.Name = if_else(Gear.Name == "Set gillnet", "SGN", Gear.Name),
         Species.Name = if_else(Species.Name == "salmon, sockeye", "Sockeye", Species.Name),
         Species.Name = if_else(Species.Name == "salmon, pink", "Pink", Species.Name),
         Harvest.Name = if_else(Harvest.Name == "State managed fishery", "CCP", Harvest.Name),
         Harvest.Name = if_else(Harvest.Name == "Private Hatchery - Fishing", "HCR", Harvest.Name)) %>% 
  unite("Year_StatWk_Gear", c(DOL.Year,Stat.Week,Gear.Name), sep = "_") %>% 
  rename(Species = Species.Name, StatArea = Stat.Area,Source = Harvest.Name) %>% 
  group_by(Year_StatWk_Gear,Species,StatArea,Source)  %>% 
  summarize(CommercialHarvest = sum(Number.Of.Animals..sum.)) %>% 
  mutate(Year_StatWk_Gear = as.factor(Year_StatWk_Gear),StatArea = as.factor(StatArea),Species = as.factor(Species),Source = as.factor(Source))

sample.df <- sample %>% 
  filter(HarvType %in% c("CCP","HCR")) %>% 
  select(-c(ID,Agency,SampleDate,SampleType,HarvLoc,District)) %>% 
  unite("Year_StatWk_Gear", c(Year,StatWk,Gear), sep = "_") %>% 
  mutate(Year_StatWk_Gear = as.factor(Year_StatWk_Gear),SubDist = as.factor(SubDist),Species = as.factor(Species),HarvType = as.factor(HarvType)) %>% 
  filter(!(HomerID == "18CPTpinks")) # Remove samples related to a Pink Salmon stomach contents pilot study

sampled.harvest <- full_join(harvest.df,sample.df,  # joining sample data frame to harvest data frame
              by = join_by(Year_StatWk_Gear == Year_StatWk_Gear, 
                         StatArea == SubDist,
                         Species == Species,
                         Source == HarvType))
```


## Change Sample Area
Some samples were identified as having a mistake made in the recording of their SubDistrict (StatArea) on the ADFG sampling form. These mistakes are rectified using the ```ChangeStatArea.csv``` file to make corrections to the ```harvest``` R datatable.

```{r changestatarea}
sampled.harvest.CHANGESTATAREA <- sampled.harvest %>% 
  filter(!(HomerID %in% c("18GPG03S","18GPG06S","18GPG11S","18PSD03S","19GSD086P","20GBT063S","21GMX066S","21GMX067P") & is.na(CommercialHarvest))) %>% 
  mutate(HomerID = ifelse(Year_StatWk_Gear == "2019_31_SGN" & Species == "Pink","19GSD086P",HomerID),
         HomerID = ifelse(Year_StatWk_Gear == "2018_24_SGN" & Species == "Sockeye" & StatArea == "24130","18GPG03S",HomerID),
         HomerID = ifelse(Year_StatWk_Gear == "2018_25_SGN" & Species == "Sockeye" & StatArea == "24130","18GPG06S",HomerID),
         HomerID = ifelse(Year_StatWk_Gear == "2018_26_SGN" & Species == "Sockeye" & StatArea == "24130","18GPG11S",HomerID),
         HomerID = ifelse(Year_StatWk_Gear == "2018_25_PS" & Species == "Sockeye" & StatArea == "24190","18PSD03S",HomerID),
         HomerID = ifelse(Year_StatWk_Gear == "2020_31_SGN" & Species == "Sockeye" & StatArea == "24106","20GBT063S",HomerID)) %>% 
  mutate(StatArea = ifelse(HomerID == "18GBT02S" & is.na(CommercialHarvest),"24117",as.character(StatArea)),
         StatArea = ifelse(HomerID == "18PSD06S" & is.na(CommercialHarvest),"24190",StatArea),
         StatArea = ifelse(HomerID == "20PSD041S" & is.na(CommercialHarvest),"24191",StatArea),
         CommercialHarvest = ifelse(HomerID == "18GBT02S" & is.na(CommercialHarvest),250,CommercialHarvest),
         CommercialHarvest = ifelse(HomerID == "18PSD06S" & is.na(CommercialHarvest),5160,CommercialHarvest),
         CommercialHarvest = ifelse(HomerID == "20PSD041S" & is.na(CommercialHarvest),3693,CommercialHarvest))
```


## Missing Harvest

Other samples where a HomerID exists (strata was sampled by ADFG staff) but there is not a corresponding commercial harvest harvest record from that stratum (subdistrict, year, week, gear-type). The harvest that the sample was taken from was reported in a different subdistrict. Thus, we must add a catch record for the strata that was sampled and then subtract a corresponding catch record from a neighboring subdistrict to keep the final catch total consistent.

### Add Harvest

First determine the sample ID's when CommercialHarvest wasn't reported and determine how many samples were taken for each ID:

```{r addmissingharvest}
missing.harvest.ID <- filter(sampled.harvest.CHANGESTATAREA, is.na(CommercialHarvest))$HomerID

pink.missing.harvest <- pink.clean %>% 
  filter(HomerSampleID %in% missing.harvest.ID) %>% 
  group_by(HomerSampleID) %>% 
  summarise(MissingHarvest = Preps)

reds.missing.harvest <- reds.clean %>% 
  filter(HomerSampleID %in% missing.harvest.ID) %>% 
  group_by(HomerSampleID) %>% 
  summarise(MissingHarvest = Preps)

add.missing.harvest <- full_join(pink.missing.harvest,reds.missing.harvest)
```

### Subtract Harvest

Harvest was added to strata that were sampled but no commercial catch was reported. However, we need to subtract the same total from the harvest of a neighboring subdistrict in the same strata (year, week, gear-type). 

``` {r subtractmissingharvest}

# Manually assign which statistical area (SubDist) that the added harvest for a sample ID (HomerID) should be subtracted from based upon our knowledge of where a catch was likely to have been reported from:
sample.missing <- filter(sample, HomerID %in% missing.harvest.ID) %>% 
  mutate(SubStatArea = ifelse(HomerID == "18PSD13S",24190,NA),
         SubStatArea = ifelse(HomerID == "18PSD25S",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "18PSD26P",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "18PSD37P",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "18PSD43P",24115,SubStatArea),
         SubStatArea = ifelse(HomerID == "19PSD088S",24191,SubStatArea),
         SubStatArea = ifelse(HomerID == "19PSD089P",24191,SubStatArea),
         SubStatArea = ifelse(HomerID == "20PSD012S",24190,SubStatArea),
         SubStatArea = ifelse(HomerID == "21PSD062P",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "21PSD068S",24190,SubStatArea),
         SubStatArea = ifelse(HomerID == "21PSD069P",24190,SubStatArea),
         SubStatArea = ifelse(HomerID == "21PSD071P",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "22PSD009S",24191,SubStatArea),
         SubStatArea = ifelse(HomerID == "22PSD039S",24193,SubStatArea),
         SubStatArea = ifelse(HomerID == "22PSD040P",24193,SubStatArea))


## Create data frame that has the final commercial harvest value for the strata that are reduced to account for allocating the sample from unreported strata   
subtract.missing.harvest <- harvest %>% 
  select(-c(ID,DOL.Month,District,Subdistrict,Stat.Area.Name,Landed.Weight..sum.,Business.Name)) %>% 
  filter(Species.Name %in% c("salmon, sockeye","salmon, pink"),Gear.Name %in% c("Purse seine","Set gillnet"),Harvest.Name %in% c("State managed fishery","Private Hatchery - Fishing")) %>% 
  mutate(Gear.Name = if_else(Gear.Name == "Purse seine", "PS", Gear.Name),
         Gear.Name = if_else(Gear.Name == "Set gillnet", "SGN", Gear.Name),
         Species.Name = if_else(Species.Name == "salmon, sockeye", "Sockeye", Species.Name),
         Species.Name = if_else(Species.Name == "salmon, pink", "Pink", Species.Name),
         Harvest.Name = if_else(Harvest.Name == "State managed fishery", "CCP", Harvest.Name),
         Harvest.Name = if_else(Harvest.Name == "Private Hatchery - Fishing", "HCR", Harvest.Name)) %>% 
  filter(Gear.Name != "SGN") %>% 
  group_by(DOL.Year,Stat.Week,Gear.Name,Species.Name,Harvest.Name,Date.of.Landing,Stat.Area)  %>% 
  summarize(CommercialHarvest = sum(Number.Of.Animals..sum.)) %>% 
  right_join(sample.missing,
             by = join_by(Species.Name == Species,
                          Stat.Area == SubStatArea,
                          Date.of.Landing == SampleDate,
                          Harvest.Name == HarvType)) %>% 
  full_join(add.missing.harvest,
            by = join_by(HomerID == HomerSampleID)) %>%
  mutate(Final.CommercialHarvest = CommercialHarvest - MissingHarvest) %>%
  select(DOL.Year,Stat.Week,Gear.Name,Species.Name,Harvest.Name,Date.of.Landing,Stat.Area,Final.CommercialHarvest)

## Recreate harvest data frame. Join with the data frame above (subtract.missing.harvest) that accounts for the reallocation of harvest to unreported strata
harvest.df <- harvest %>% 
  select(-c(ID,DOL.Month,District,Subdistrict,Stat.Area.Name,Landed.Weight..sum.,Business.Name)) %>% 
  filter(Species.Name %in% c("salmon, sockeye","salmon, pink"),Gear.Name %in% c("Purse seine","Set gillnet"),Harvest.Name %in% c("State managed fishery","Private Hatchery - Fishing")) %>% 
  mutate(Gear.Name = if_else(Gear.Name == "Purse seine", "PS", Gear.Name),
         Gear.Name = if_else(Gear.Name == "Set gillnet", "SGN", Gear.Name),
         Species.Name = if_else(Species.Name == "salmon, sockeye", "Sockeye", Species.Name),
         Species.Name = if_else(Species.Name == "salmon, pink", "Pink", Species.Name),
         Harvest.Name = if_else(Harvest.Name == "State managed fishery", "CCP", Harvest.Name),
         Harvest.Name = if_else(Harvest.Name == "Private Hatchery - Fishing", "HCR", Harvest.Name)) %>% 
  group_by(DOL.Year,Stat.Week,Gear.Name,Species.Name,Harvest.Name,Date.of.Landing,Stat.Area)  %>% 
  summarize(CommercialHarvest = sum(Number.Of.Animals..sum.)) %>% 
  full_join(subtract.missing.harvest,
            by = join_by(DOL.Year, Stat.Week, Gear.Name, Species.Name, Harvest.Name, Date.of.Landing, Stat.Area)) %>% 
  mutate(CommercialHarvest = ifelse(!is.na(Final.CommercialHarvest),Final.CommercialHarvest,CommercialHarvest)) %>% 
  select(-Final.CommercialHarvest) %>% 
  unite("Year_StatWk_Gear", c(DOL.Year,Stat.Week,Gear.Name), sep = "_") %>% 
  rename(Species = Species.Name, StatArea = Stat.Area, Source = Harvest.Name) %>% 
  group_by(Year_StatWk_Gear,Species,StatArea,Source) %>% 
  summarise(CommercialHarvest = sum(CommercialHarvest)) %>% 
  mutate(Year_StatWk_Gear = as.factor(Year_StatWk_Gear),StatArea = as.factor(StatArea),Species = as.factor(Species),Source = as.factor(Source))

## Rejoin harvest dataframe with sample dataframe
sampled.harvest.SUBTRACTED <- full_join(harvest.df,sampled.harvest.CHANGESTATAREA,  # joining sample data frame to harvest data frame
              by = join_by(Year_StatWk_Gear == Year_StatWk_Gear, 
                         StatArea == StatArea,
                         Species == Species,
                         Source == Source)) %>% 
  mutate(CommercialHarvest = CommercialHarvest.x) %>% 
  select(-c(CommercialHarvest.x,CommercialHarvest.y))

## Add in samples for strata that were sampled but no catch was reported:
sampled.harvest.ADDED <- sampled.harvest.SUBTRACTED %>% 
  full_join(add.missing.harvest, by = join_by(HomerID == HomerSampleID)) %>% 
  mutate(CommercialHarvest = ifelse(is.na(CommercialHarvest),MissingHarvest,CommercialHarvest)) %>% 
  select(-MissingHarvest) %>% 
  mutate(CommercialHarvest = ifelse(Year_StatWk_Gear == "2021_31_PS" & Species == "Pink" & StatArea == 24191,55+76,CommercialHarvest)) ## Individual edit of data where two HomerID samples were added to the same strata commercial harvest

```

## Sampled Harvest Data
``` {r cleansampledharvest}
sampled.harvest.CLEAN <- sampled.harvest.ADDED %>% # Rename sampled.harvest data frame to more intuitive name
  mutate(StatArea = as.factor(StatArea)) %>% 
  filter(str_detect(StatArea,"^241"))
```

# Analysis with StatWk part of Stratum
# 1 - Estimate of C.hat~hi~

The otolith-derived estimate of the contribution of hatchery-area *h* to district-period stratum *i*, *C~hi~* is
$$\hat{C}_{hi} = \frac{o_{hi}}{n_i}N_i$$
where,

*h* = Hatchery-area (one of LCI, PWS, KOD, or Other)

*o~hi~* = Number of otoliths from hatchery district *h* in sample *n~i~*,

*n~i~* = Number of otoliths sampled from stratum *i* (usually 96), and

*N~i~* = Number of fish caught in stratum *i*.

The variance estimate of *C~hi~* is
$$var(\hat{C}_{hi}) = N_i^2\frac{1}{n_i-1}\frac{o_{hi}}{n_i}(1-\frac{o_{hi}}{n_i})$$

```{r C.hat_hi, message=FALSE}

# Create table of # of otoliths from hatchery-area (h) in sample from a given stratum (n_i):
sample.o_hi <- otolith.TM.data %>% 
  pivot_longer(cols = c("LCI","PWS","KOD","Other"), names_to = "HatcheryArea", values_to = "MarkedOtoliths") %>% # Pivot table to LONG format, with one count of the # of MarkedOtoliths per Strata*HatcheryArea combination
  rename(HomerID = HomerSampleID) %>% 
  mutate(HatcheryArea = as.factor(HatcheryArea)) %>% 
  group_by(Year_StatWk_Gear,Species,Source,StatArea,HatcheryArea) %>% 
  summarise(o_hi = sum(MarkedOtoliths))

# Create table of total # of otoliths sampled from a given stratum (n_i):
sample.n_i <- otolith.TM.data %>% 
  group_by(Year_StatWk_Gear,Species,Source,StatArea) %>% # Group into stratum (i)
  summarise(n_i = sum(NotMarked) + sum(Marked)) # SampledOtoliths = total # of otoliths sampled

# Create table of total # of fish commercially harvested in stratum (i):
harvest.N_i <- sampled.harvest.CLEAN %>%
  group_by(Year_StatWk_Gear,Species,Source,StatArea) %>% 
  summarise(N_i = unique(CommercialHarvest))

# Create a table of the otolith-derived estimated of the contribution of hatchery-area (h) to period-gear-district stratum (i):
C.hat_hi <- sample.o_hi %>% 
  full_join(sample.n_i,
            by = join_by(Year_StatWk_Gear,Species,Source,StatArea)) %>% 
  full_join(harvest.N_i,
            by = join_by(Year_StatWk_Gear,Species,Source,StatArea)) %>% 
  filter(!is.na(HatcheryArea)) %>% # Removing instances where HatcheryArea = NA, aka when a stratum had commercial harvest occur but no sampling occurred
  mutate(ProportionMarked = round(o_hi/n_i, 3), # Calculate hatchery proportion of sample
         C.hat_hi = round(ProportionMarked*N_i,1),# EQUATION 1: Multiply hatchery proportion by the total commercial harvest for estimate hatchery proportion in commercial catch
         C.hat_hi_VAR = as.integer((N_i^2)*(1/(n_i-1))*(o_hi/n_i)*(1 - (o_hi/n_i)))) # EQUATION 2: Variance estimate of the hatchery-derived commercial harvest contribution

```

```{r C.hat_hi.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_hi, desc(C.hat_hi)),10))
```

# 2 - Estimate of C.hat~Sh~

Otolith-derived estimates of the contribution of hatchery-area *h*, *C~Sh~*, to all sampled harvests (in a given StatWk but combined across all StatArea's) will be calculated as
$$\hat{C}_{Sh}=\sum_{i=1}^{Q}C_{hi}$$
where,

*Q* = Number of recovery strata in which otoliths from hatchery-area *h* are found.

The variance estimate of *C.hat~Sh~* is
$$var(\hat{C}_{Sh})=\sum_{i=1}^{Q}var(C_{hi})$$

```{r C.hat_Sh}
C.hat_Sh <- C.hat_hi %>% 
  group_by(Year_StatWk_Gear,Species,Source,HatcheryArea) %>% 
  summarize(C.hat_Sh = sum(C.hat_hi), # EQUATION 3: Estimate of contribution of hatchery-area (h) to all sampled harvests within a StatWk, combined across all sampled StatArea's.
            C.hat_Sh_VAR = sum(C.hat_hi_VAR)) # EQUATION 4: Variance estimate of C.hat_Sh
```

```{r C.hat_Sh.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_Sh, desc(C.hat_Sh)),10))
```

# 3 - Estimate of C.hat~Uh~

The contribution of hatchery-area *h* to unsampled strata, *C~Uh~*, will be estimated from contribution rates associated with strata sampled from the same district-period openings as the unsampled strata using methods similar to those used for coded wire tags (Riffe et al. 1996):
$$\hat{C}_{Uh}=\sum_{i=1}^{U}\left[N_i\left(\frac{\sum_{j=1}^{S}\hat{C}_{hj}}{\sum_{j=1}^{S}N_j}\right)\right]$$
where,

*U* = Number of unsampled strata,

*N~i~* = Number of fish in *i*^th^ unsampled stratum,

*S* = Number of strata sampled in the period in which the unsampled stratum resides,

*C~hj~* = Contribution of thermal mark *h* to the sampled stratum *j*, and

*N~j~* = Number of fish in *j*^th^ sampled stratum.

The variance estimate of *C.hat~Uh~* is

$$var(\hat{C}_{Uh})=\sum_{i=1}^{U}\left[N_{i}^2\left(\frac{1}{\sum_{j=1}^{S}N_{j}-1}\right)\left(\frac{\sum_{j=1}^{S}\hat{C}_{hj}}{\sum_{j=1}^{S}N_j}\right)\left(1-\frac{\sum_{j=1}^{S}\hat{C}_{hj}}{\sum_{j=1}^{S}N_j}\right)\right]$$

```{r C.hat_Uh}
unsampled.N_i <- sample.o_hi %>%
  full_join(sample.n_i,
            by = join_by(Year_StatWk_Gear,Species,Source,StatArea)) %>%
  full_join(harvest.N_i,
            by = join_by(Year_StatWk_Gear,Species,Source,StatArea)) %>%
  filter(is.na(HatcheryArea)) %>% # Selecting instances where HatcheryArea = NA, aka when a stratum had commercial harvest occur but no sampling occurred
  select(-c(HatcheryArea,o_hi,n_i))

unsampled.C.hat_Uh <- C.hat_hi %>% 
  group_by(Year_StatWk_Gear,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_hj = sum(C.hat_hi), N_j = sum(N_i),ProportionMarked = C.hat_hj/N_j)

C.hat_Uh <- unsampled.N_i %>% 
  left_join(unsampled.C.hat_Uh,
            by = join_by(Year_StatWk_Gear,Species,Source),
            relationship = "many-to-many") %>% 
  group_by(Year_StatWk_Gear,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_Uh = round(sum(N_i)*unique(ProportionMarked),1), # EQUATION 5: Estimate the contribution of hatchery-area (h) to unsampled StatArea's
            C.hat_Uh_VAR = as.integer(sum(N_i^2 * (1/(sum(N_j)-1)) * (sum(C.hat_hj)/sum(N_j)) * (1 - (sum(C.hat_hj)/sum(N_j)))   ))) %>% # EQUATION 6: Variance estimate of C.hat_Uh
  filter(!is.na(HatcheryArea))

# NA's occur in the above data frame (C.hat_Uh) when there is commercial harvest from a stratum (Year_StatWk_Gear,Species,Fishery) but no fish were sampled in that stratum (across any StatArea's) from which to estimate hatchery proportion...
# These unsampled stratum observations are removed from the C.hat_Uh data frame...

# And these unsampled stratum observations are subsetted to their own data frame: 
unsampled.stratum <- unsampled.N_i %>% 
  left_join(unsampled.C.hat_Uh,
            by = join_by(Year_StatWk_Gear,Species,Source),
            relationship = "many-to-many") %>% 
  group_by(Year_StatWk_Gear,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_Uh = sum(N_i)*unique(ProportionMarked),
            C.hat_Uh_VAR = sum(N_i^2 * (1/(sum(N_j)-1)) * (sum(C.hat_hj)/sum(N_j)) * (1 - (sum(C.hat_hj)/sum(N_j)))   )) %>% 
  filter(is.na(HatcheryArea)) %>% 
  select(Year_StatWk_Gear,Species,Source) %>% 
  unite("ID",c(Year_StatWk_Gear,Species,Source), sep = "_")
# And the hatchery contribution may be estimated to these observations in Part 6 -- maybe...

```

```{r C.hat_Uh.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_Uh, desc(C.hat_Uh)),10))
```

# 4 - Estimate of C.hat~h~

An estimate of the contribution by hatchery-area *h* to all strata, sampled and unsampled, will be calculated by
$$\hat{C}_h=\hat{C}_{Sh}+\hat{C}_{Uh}$$

The variance estimate of C.hat~h~ is
$$var(\hat{C}_h)=var(\hat{C}_{Sh})+var(\hat{C}_{Uh})$$

```{r C.hat_h}
C.hat_h <- C.hat_Sh %>% 
  full_join(C.hat_Uh,
            by = join_by(Year_StatWk_Gear,Species,Source,HatcheryArea)) %>% 
  replace_na(list(C.hat_Uh = 0, C.hat_Uh_VAR = 0)) %>% 
  mutate(C.hat_h = C.hat_Sh + C.hat_Uh, # EQUATION 7: Estimate of contribution of hatchery-area (h) to all strata
         C.hat_h_VAR = C.hat_Sh_VAR + C.hat_Uh_VAR) # EQUATION 8: Variance estimate of C.hat_h
```

```{r C.hat_h.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_h, desc(C.hat_h)),10))
```

# 5 - Estimate of C.hat~H~

The overall hatchery contribution by all hatcheries is calculated by

$$\hat{C}_H=\sum_{h}^{H}\hat{C}_h$$
```{r C.hat_H}
C.hat_H <- C.hat_h %>% 
  separate(Year_StatWk_Gear, into = c("Year","StatWk","Gear"), sep = "_") %>% 
  group_by(Gear,Year,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_H = sum(C.hat_h),
            C.hat_H_VAR = sum(C.hat_h_VAR)) %>% 
  mutate(C.hat_H_se = as.integer(sqrt(C.hat_H_VAR)),
         lower95CI = C.hat_H - 1.96*C.hat_H_se,
         upper95CI = C.hat_H + 1.96*C.hat_H_se)

C.hat_H.TOTAL <- C.hat_h %>% 
  separate(Year_StatWk_Gear, into = c("Year","StatWk","Gear"), sep = "_") %>% 
  group_by(Gear,Year,Species,Source) %>% 
  summarise(C.hat_H = sum(C.hat_h),
            C.hat_H_VAR = sum(C.hat_h_VAR)) %>% 
  mutate(C.hat_H_se = as.integer(sqrt(C.hat_H_VAR)),
         lower95CI = C.hat_H - 1.96*C.hat_H_se,
         upper95CI = C.hat_H + 1.96*C.hat_H_se)
```

```{r C.hat_H.table, echo=FALSE}
knitr::kable(C.hat_H)
```

# Catch Proportion Comparison

```{r catch.comparison}
catch.comparison.HATCHAREA <- harvest.N_i %>% 
  unite("ID",c(Year_StatWk_Gear,Species,Source), sep = "_") %>% 
  filter(!(ID %in% unsampled.stratum$ID)) %>% 
  separate(ID, into = c("Year","StatWk","Gear","Species","Source"), sep = "_") %>% 
  group_by(Gear,Year,Species,Source) %>% 
  summarize(Catch = sum(N_i)) %>% 
  left_join(C.hat_H,
            by = join_by(Gear,Year,Species,Source)) %>% 
  mutate(HatcheryProportion = round(C.hat_H/Catch,3))

catch.comparison.TOTAL <- harvest.N_i %>% 
  unite("ID",c(Year_StatWk_Gear,Species,Source), sep = "_") %>% 
  filter(!(ID %in% unsampled.stratum$ID)) %>% 
  separate(ID, into = c("Year","StatWk","Gear","Species","Source"), sep = "_") %>% 
  group_by(Gear,Year,Species,Source) %>% 
  summarize(Catch = sum(N_i)) %>% 
  left_join(C.hat_H.TOTAL,
            by = join_by(Gear,Year,Species,Source)) %>% 
  mutate(HatcheryProportion = round(C.hat_H/Catch,3))
```

```{r catchcomp.table, echo=FALSE}
knitr::kable(catch.comparison)
```

# Analysis w/out StatWk part of Stratum

```{r restructuredata}
otolith.TM.data2 <- otolith.TM.data %>% 
  separate(Year_StatWk_Gear, into = c("Year","StatWk","Gear"), sep = "_")

sampled.harvest.CLEAN2 <- sampled.harvest.CLEAN %>% 
  separate(Year_StatWk_Gear, into = c("Year","StatWk","Gear"), sep = "_")

```

# 1 - Estimate of C.hat~hi~

The otolith-derived estimate of the contribution of hatchery-area *h* to district-period stratum *i*, *C~hi~* is
$$\hat{C}_{hi} = \frac{o_{hi}}{n_i}N_i$$
where,

*h* = Hatchery-area (one of LCI, PWS, KOD, or Other)

*o~hi~* = Number of otoliths from hatchery district *h* in sample *n~i~*,

*n~i~* = Number of otoliths sampled from stratum *i* (usually 96), and

*N~i~* = Number of fish caught in stratum *i*.

The variance estimate of *C~hi~* is
$$var(\hat{C}_{hi}) = N_i^2\frac{1}{n_i-1}\frac{o_{hi}}{n_i}(1-\frac{o_{hi}}{n_i})$$

```{r C.hat_hi, message=FALSE}

# Create table of # of otoliths from hatchery-area (h) in sample from a given stratum (n_i):
sample.o_hi2 <- otolith.TM.data2 %>% 
  pivot_longer(cols = c("LCI","PWS","KOD","Other"), names_to = "HatcheryArea", values_to = "MarkedOtoliths") %>% # Pivot table to LONG format, with one count of the # of MarkedOtoliths per Strata*HatcheryArea combination
  rename(HomerID = HomerSampleID) %>% 
  mutate(HatcheryArea = as.factor(HatcheryArea)) %>% 
  group_by(Year,Gear,Species,Source,StatArea,HatcheryArea) %>% 
  summarise(o_hi = sum(MarkedOtoliths))

# Create table of total # of otoliths sampled from a given stratum (n_i):
sample.n_i2 <- otolith.TM.data2 %>% 
  group_by(Year,Gear,Species,Source,StatArea) %>% # Group into stratum (i)
  summarise(n_i = sum(NotMarked) + sum(Marked)) # SampledOtoliths = total # of otoliths sampled

# Create table of total # of fish commercially harvested in stratum (i):
harvest.N_i2 <- sampled.harvest.CLEAN2 %>%
  group_by(Year,Gear,Species,Source,StatArea) %>% 
  summarise(N_i = sum(unique(CommercialHarvest)))

# Create a table of the otolith-derived estimated of the contribution of hatchery-area (h) to period-gear-district stratum (i):
C.hat_hi2 <- sample.o_hi2 %>% 
  full_join(sample.n_i2,
            by = join_by(Year,Gear,Species,Source,StatArea)) %>% 
  full_join(harvest.N_i2,
            by = join_by(Year,Gear,Species,Source,StatArea)) %>% 
  filter(!is.na(HatcheryArea)) %>% # Removing instances where HatcheryArea = NA, aka when a stratum had commercial harvest occur but no sampling occurred
  mutate(ProportionMarked = o_hi/n_i, # Calculate hatchery proportion of sample
         C.hat_hi = ProportionMarked*N_i,# EQUATION 1: Multiply hatchery proportion by the total commercial harvest for estimate hatchery proportion in commercial catch
         C.hat_hi_VAR = (N_i^2)*(1/(n_i-1))*(o_hi/n_i)*(1 - (o_hi/n_i))) # EQUATION 2: Variance estimate of the hatchery-derived commercial harvest contribution

```

```{r C.hat_hi.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_hi, desc(C.hat_hi)),10))
```

# 2 - Estimate of C.hat~Sh~

Otolith-derived estimates of the contribution of hatchery-area *h*, *C~Sh~*, to all sampled harvests (in a given StatWk but combined across all StatArea's) will be calculated as
$$\hat{C}_{Sh}=\sum_{i=1}^{Q}C_{hi}$$
where,

*Q* = Number of recovery strata in which otoliths from hatchery-area *h* are found.

The variance estimate of *C.hat~Sh~* is
$$var(\hat{C}_{Sh})=\sum_{i=1}^{Q}var(C_{hi})$$

```{r C.hat_Sh}
C.hat_Sh2 <- C.hat_hi2 %>% 
  group_by(Year,Gear,Species,Source,HatcheryArea) %>% 
  summarize(C.hat_Sh = sum(C.hat_hi), # EQUATION 3: Estimate of contribution of hatchery-area (h) to all sampled harvests within a StatWk, combined across all sampled StatArea's.
            C.hat_Sh_VAR = sum(C.hat_hi_VAR)) # EQUATION 4: Variance estimate of C.hat_Sh
```

```{r C.hat_Sh.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_Sh, desc(C.hat_Sh)),10))
```

# 3 - Estimate of C.hat~Uh~

The contribution of hatchery-area *h* to unsampled strata, *C~Uh~*, will be estimated from contribution rates associated with strata sampled from the same district-period openings as the unsampled strata using methods similar to those used for coded wire tags (Riffe et al. 1996):
$$\hat{C}_{Uh}=\sum_{i=1}^{U}\left[N_i\left(\frac{\sum_{j=1}^{S}\hat{C}_{hj}}{\sum_{j=1}^{S}N_j}\right)\right]$$
where,

*U* = Number of unsampled strata,

*N~i~* = Number of fish in *i*^th^ unsampled stratum,

*S* = Number of strata sampled in the period in which the unsampled stratum resides,

*C~hj~* = Contribution of thermal mark *h* to the sampled stratum *j*, and

*N~j~* = Number of fish in *j*^th^ sampled stratum.

The variance estimate of *C.hat~Uh~* is

$$var(\hat{C}_{Uh})=\sum_{i=1}^{U}\left[N_{i}^2\left(\frac{1}{\sum_{j=1}^{S}N_{j}-1}\right)\left(\frac{\sum_{j=1}^{S}\hat{C}_{hj}}{\sum_{j=1}^{S}N_j}\right)\left(1-\frac{\sum_{j=1}^{S}\hat{C}_{hj}}{\sum_{j=1}^{S}N_j}\right)\right]$$

```{r C.hat_Uh}
unsampled.N_i2 <- sample.o_hi2 %>%
  full_join(sample.n_i2,
            by = join_by(Year,Gear,Species,Source,StatArea)) %>%
  full_join(harvest.N_i2,
            by = join_by(Year,Gear,Species,Source,StatArea)) %>%
  filter(is.na(HatcheryArea)) %>% # Selecting instances where HatcheryArea = NA, aka when a stratum had commercial harvest occur but no sampling occurred
  select(-c(HatcheryArea,o_hi,n_i))

unsampled.C.hat_Uh2 <- C.hat_hi2 %>% 
  group_by(Year,Gear,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_hj = sum(C.hat_hi), N_j = sum(N_i),ProportionMarked = C.hat_hj/N_j)

C.hat_Uh2 <- unsampled.N_i2 %>% 
  left_join(unsampled.C.hat_Uh2,
            by = join_by(Year,Gear,Species,Source),
            relationship = "many-to-many") %>% 
  group_by(Year,Gear,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_Uh = sum(N_i)*unique(ProportionMarked), # EQUATION 5: Estimate the contribution of hatchery-area (h) to unsampled StatArea's
            C.hat_Uh_VAR = as.integer(sum(N_i^2 * (1/(sum(N_j)-1)) * (sum(C.hat_hj)/sum(N_j)) * (1 - (sum(C.hat_hj)/sum(N_j)))   ))) %>% # EQUATION 6: Variance estimate of C.hat_Uh
  filter(!is.na(HatcheryArea))

# NA's occur in the above data frame (C.hat_Uh) when there is commercial harvest from a stratum (Year_StatWk_Gear,Species,Fishery) but no fish were sampled in that stratum (across any StatArea's) from which to estimate hatchery proportion...
# These unsampled stratum observations are removed from the C.hat_Uh data frame...

# And these unsampled stratum observations are subsetted to their own data frame: 
unsampled.stratum2 <- unsampled.N_i2 %>% 
  left_join(unsampled.C.hat_Uh2,
            by = join_by(Year,Gear,Species,Source),
            relationship = "many-to-many") %>% 
  group_by(Year,Gear,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_Uh = sum(N_i)*unique(ProportionMarked),
            C.hat_Uh_VAR = sum(N_i^2 * (1/(sum(N_j)-1)) * (sum(C.hat_hj)/sum(N_j)) * (1 - (sum(C.hat_hj)/sum(N_j)))   )) %>% 
  filter(is.na(HatcheryArea)) %>% 
  select(Year,Gear,Species,Source) %>% 
  unite("ID",c(Year,Gear,Species,Source), sep = "_")
# And the hatchery contribution may be estimated to these observations in Part 6 -- maybe...

```

```{r C.hat_Uh.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_Uh, desc(C.hat_Uh)),10))
```

# 4 - Estimate of C.hat~h~

An estimate of the contribution by hatchery-area *h* to all strata, sampled and unsampled, will be calculated by
$$\hat{C}_h=\hat{C}_{Sh}+\hat{C}_{Uh}$$

The variance estimate of C.hat~h~ is
$$var(\hat{C}_h)=var(\hat{C}_{Sh})+var(\hat{C}_{Uh})$$

```{r C.hat_h}
C.hat_h2 <- C.hat_Sh2 %>% 
  full_join(C.hat_Uh2,
            by = join_by(Year,Gear,Species,Source,HatcheryArea)) %>% 
  replace_na(list(C.hat_Uh = 0, C.hat_Uh_VAR = 0)) %>% 
  mutate(C.hat_h = C.hat_Sh + C.hat_Uh, # EQUATION 7: Estimate of contribution of hatchery-area (h) to all strata
         C.hat_h_VAR = C.hat_Sh_VAR + C.hat_Uh_VAR) # EQUATION 8: Variance estimate of C.hat_h
```

```{r C.hat_h.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_h, desc(C.hat_h)),10))
```

# 5 - Estimate of C.hat~H~

The overall hatchery contribution by all hatcheries is calculated by

$$\hat{C}_H=\sum_{h}^{H}\hat{C}_h$$
```{r C.hat_H}
C.hat_H2 <- C.hat_h2 %>% 
  # separate(Year_StatWk_Gear, into = c("Year","StatWk","Gear"), sep = "_") %>% 
  group_by(Gear,Year,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_H = sum(C.hat_h),
            C.hat_H_VAR = sum(C.hat_h_VAR)) %>% 
  mutate(C.hat_H_se = as.integer(sqrt(C.hat_H_VAR)),
         lower95CI = C.hat_H - 1.96*C.hat_H_se,
         upper95CI = C.hat_H + 1.96*C.hat_H_se)

C.hat_H.TOTAL2 <- C.hat_h2 %>% 
  # separate(Year_StatWk_Gear, into = c("Year","StatWk","Gear"), sep = "_") %>% 
  group_by(Gear,Year,Species,Source) %>% 
  summarise(C.hat_H = sum(C.hat_h),
            C.hat_H_VAR = sum(C.hat_h_VAR)) %>% 
  mutate(C.hat_H_se = as.integer(sqrt(C.hat_H_VAR)),
         lower95CI = C.hat_H - 1.96*C.hat_H_se,
         upper95CI = C.hat_H + 1.96*C.hat_H_se)
```

```{r C.hat_H.table, echo=FALSE}
knitr::kable(C.hat_H)
```

# Catch Proportion Comparison

```{r catch.comparison}
catch.comparison.HATCHAREA2 <- harvest.N_i2 %>% 
  unite("ID",c(Year,Gear,Species,Source), sep = "_") %>% 
  filter(!(ID %in% unsampled.stratum2$ID)) %>% 
  separate(ID, into = c("Year","Gear","Species","Source"), sep = "_") %>% 
  group_by(Gear,Year,Species,Source) %>% 
  summarize(Catch = sum(N_i)) %>% 
  left_join(C.hat_H2,
            by = join_by(Gear,Year,Species,Source)) %>% 
  mutate(HatcheryProportion = C.hat_H/Catch)

catch.comparison.TOTAL2 <- harvest.N_i2 %>% 
  unite("ID",c(Year,Gear,Species,Source), sep = "_") %>% 
  filter(!(ID %in% unsampled.stratum2$ID)) %>% 
  separate(ID, into = c("Year","Gear","Species","Source"), sep = "_") %>% 
  group_by(Gear,Year,Species,Source) %>% 
  summarize(Catch = sum(N_i)) %>% 
  left_join(C.hat_H.TOTAL2,
            by = join_by(Gear,Year,Species,Source)) %>% 
  mutate(HatcheryProportion = C.hat_H/Catch)
```

```{r catchcomp.table, echo=FALSE}
knitr::kable(catch.comparison)
```

# 6 - C.hat~H~ in Unsampled Strata
Occasionally, a given strata (Year, StatWk, Gear, StatArea) did not have a sample taken. In these instances, the hatchery contribution was estimated as the average hatchery-origin proportion across all StatWk's for a given StatArea multiplied by the harvest occurring in the unsampled strata:

``` {r pink.unsampledstrata}
pink.stratum.TM <- pink.C.hat %>% 
  separate(Year_StatWk_Gear, into = c("Year","StatWk","Gear")) %>% 
  group_by(Year,Gear,StatWk,StatArea) %>% 
  summarise(Marked = sum(MarkedOtoliths),Sampled = unique(SampledOtoliths))

pink.est.unsampled.stratum <- pink.unsampled.stratum %>% 
  left_join(filter(sampled.harvest.CLEAN,Species == "Pink" & str_detect(StatArea, "^241"))) %>% 
  filter(!is.na(CommercialHarvest)) %>% 
  separate(Year_StatWk_Gear, into = c("Year","StatWk","Gear")) %>% 
  select(-StatWk) %>% 
  left_join(pink.stratum.TM) %>% 
  group_by(Year,Gear,StatArea,CommercialHarvest) %>% 
  summarise(Marked = sum(Marked),Sampled = sum(Sampled)) %>% 
  mutate(Proportion = Marked/Sampled, Contribution = CommercialHarvest*Proportion)

pink.C.hat_unsampled_Year.StatWk.Gear.StatArea <- pink.est.unsampled.stratum %>% 
  filter(!is.na(Contribution)) %>% 
  group_by(Year,Gear) %>% 
  summarise(Addl_C.hat1 = sum(Contribution))
```

In other instances, a StatArea had harvest that occurred but did not have a sample analyzed for any StatWk's. In these instances, the hatchery contribution was estimated as the overall hatchery proportion across all StatArea's and StatWk's in the same Year and Gear strata multiplied by the unsampled commercial harvest: 

``` {r pink.unsampledstrata2}
pink.C.hat_unsampled_Year.Gear.StatArea <- pink.est.unsampled.stratum %>% 
  filter(is.na(Contribution)) %>% 
  group_by(Year,Gear) %>% 
  summarise(CommercialHarvest = sum(CommercialHarvest)) %>% 
  left_join(pink.catch.comparison) %>% 
  mutate(Addl_C.hat2 = HatcheryProportion*CommercialHarvest)
```

Finally, the changes to the overall hatchery contribution across all hatcheries and all sampled/unsampled harvest is:

```{r pink.FINAL}
pink.C.hat_H.FINAL <- pink.catch.comparison %>% 
  left_join(pink.C.hat_unsampled_Year.StatWk.Gear.StatArea) %>% 
  left_join(pink.C.hat_unsampled_Year.Gear.StatArea) %>% 
  replace_na(list(Addl_C.hat1 = 0)) %>% 
  select(-CommercialHarvest) %>% 
  mutate(Contrib1 = round(Addl_C.hat1/ReportedCatch,3), 
         Contrib2 = round(Addl_C.hat2/ReportedCatch,3),
         C.hat_H.FINAL = C.hat_H + Addl_C.hat1 + Addl_C.hat2,
         HatcheryProportion.FINAL = round(C.hat_H.FINAL/ReportedCatch,3))
```

```{r pink.6.table, echo=FALSE}
knitr::kable(pink.C.hat_H.FINAL)
```

*I am unsure how to estimate variance in this final estimate of C.hat_H (which includes sampled and unsampled harvests).*

**It is my opinion that we only estimate hatchery proportion in sampled strata and not extrapolate to harvests in unsampled strata.**

