---
title: "Analysis of LCI Thermal Mark Hatchery Otoliths, 2018-2022"
author: "Jonah Bacon"
date: "`r Sys.Date()`"
output: 
  html_document:
      toc: true
      toc_float:
        collapsed: false
        smooth_scroll: true
---

This code analyzes data collected from 2018-2022 as part of the Thermal Mark Hatchery Salmon Otoliths in Lower Cook Inlet project (*REGIONAL OPERATIONAL PLAN NO. ROP.CF.2A.2023.XX*). Data, scripts, and associated files are located in the Homer LAN network: ```O:\DCF\SALMON\OTOLITH_STUDIES\1_ANALYSES\2018-2022```. Data are first cleaned, edited (when commercial harvest was misreported or errors in recording occurred), and then analyzed for Pink Salmon followed by identical analysis for Sockeye Salmon.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "O:/DCF/JonahB/R Projects/OtolithTM_analysis/2018-2022")

# Load libraries
library(tidyverse)

```

# Prepare Data
## Data

Load data files from ```O:\DCF\SALMON\OTOLITH_STUDIES\1_ANALYSES\2018-2022\input```.

```{r loaddata}
pink.marks <- read.csv("input/PinkMarks.csv")
reds.marks <- read.csv("input/RedMarks.csv")
harvest <- read.csv("input/tblHarvest.csv")
sample <- read.csv("input/tblHomerSample.csv")
ChangeStatArea <- read.csv("input/ChangeStatArea.csv")
MissingHarvest <- read.csv("input/MissingHarvest.csv")
```

## Clean Data

```{r cleanotolithdata}

pink.marks$StatArea <- as.character(pink.marks$StatArea)
pink.clean <- pink.marks %>% 
  filter(Source %in% c("COMMERCIAL COM PROP","COST RECOVERY")) %>%
  select(-c(Agency,FisheryName,Species,SubDistrict,AStreamCode,Stock,Harvest,SampleDate,SurveySite)) %>% 
  separate(StatArea, into = c("StatArea","Extra"), sep = 5) %>% 
  mutate(StatArea = as.integer(StatArea)) %>% 
  filter(StatArea > 10000) %>% 
  select(-Extra) %>% 
  mutate(# StatArea = if_else(Gear == "SET GILLNET", "24100", StatArea),
         StatArea = ifelse(HomerSampleID == "18PSD45P",24107,StatArea), # Specific correction
         StatArea = ifelse(HomerSampleID == "19GSD086P",24106,StatArea), # Specific correction, don't need any longer?
         Gear = if_else(Gear == "PURSE SEINE", "PS", Gear),
         Gear = if_else(Gear == "SET GILLNET", "SGN", Gear),
         Source = if_else(Source == "COMMERCIAL COM PROP", "CCP", Source),
         Source = if_else(Source == "COST RECOVERY", "HCR", Source)) %>% 
  mutate(Year = as.factor(Year),Gear = as.factor(Gear),StatArea = as.factor(StatArea),Source = as.factor(Source),Species = "Pink") %>% 
  rename(StatWk = StatWeek)

reds.clean <- reds.marks %>% 
  filter(Source %in% c("COMMERCIAL COM PROP","COST RECOVERY")) %>%
  select(-c(Agency,FisheryName,Species,SubDistrict,AStreamCode,Stock,Harvest,SampleDate,SurveySite)) %>% 
  separate(StatArea, into = c("StatArea","Extra"), sep = 5) %>% 
  mutate(StatArea = as.integer(StatArea)) %>% 
  filter(StatArea > 10000) %>% 
  select(-Extra) %>% 
  mutate(# StatArea = if_else(Gear == "SET GILLNET", "24100", StatArea),
         Gear = if_else(Gear == "PURSE SEINE", "PS", Gear),
         Gear = if_else(Gear == "SET GILLNET", "SGN", Gear),
         Source = if_else(Source == "COMMERCIAL COM PROP", "CCP", Source),
         Source = if_else(Source == "COST RECOVERY", "HCR", Source),
         StatArea = ifelse(HomerSampleID == "18PSD06S",24190,StatArea),
         StatArea = ifelse(HomerSampleID == "18PSD03S",24190,StatArea),
         StatArea = ifelse(HomerSampleID == "18GHC10S",24115,StatArea),
         StatArea = ifelse(HomerSampleID == "18GPG11S",24130,StatArea),
         StatArea = ifelse(HomerSampleID == "18GPG06S",24130,StatArea),
         StatArea = ifelse(HomerSampleID == "18GPG03S",24130,StatArea),
         StatArea = ifelse(HomerSampleID == "18PSD06S",24190,StatArea),
         StatArea = ifelse(HomerSampleID == "18PSD03S",24190,StatArea)) %>% 
  mutate(StatWeek = ifelse(HomerSampleID == "22GMX005S",25,StatWeek),
         Year = as.factor(Year),Gear = as.factor(Gear),StatArea = as.factor(StatArea),Source = as.factor(Source),Species = "Sockeye") %>% 
  rename(StatWk = StatWeek)

otolith.TM.data <- pink.clean %>% 
  select(HomerSampleID,Year,StatWk,Gear,Species,Source,StatArea,Preps,NotMarked,Marked,LCI,PWS,KOD,Other) %>% 
  full_join(select(reds.clean, c(HomerSampleID,Year,StatWk,Gear,Species,Source,StatArea,Preps,NotMarked,Marked,LCI,PWS,KOD,Other)))
```

```{r cleanharvestdata}
harvest.df <- harvest %>% 
  select(-c(ID,DOL.Month,District,Subdistrict,Stat.Area.Name,Landed.Weight..sum.,Business.Name)) %>% 
  filter(Species.Name %in% c("salmon, sockeye","salmon, pink"),Gear.Name %in% c("Purse seine","Set gillnet"),Harvest.Name %in% c("State managed fishery","Private Hatchery - Fishing")) %>% 
  mutate(Gear.Name = if_else(Gear.Name == "Purse seine", "PS", Gear.Name),
         Gear.Name = if_else(Gear.Name == "Set gillnet", "SGN", Gear.Name),
         Species.Name = if_else(Species.Name == "salmon, sockeye", "Sockeye", Species.Name),
         Species.Name = if_else(Species.Name == "salmon, pink", "Pink", Species.Name),
         Harvest.Name = if_else(Harvest.Name == "State managed fishery", "CCP", Harvest.Name),
         Harvest.Name = if_else(Harvest.Name == "Private Hatchery - Fishing", "HCR", Harvest.Name)) %>% 
  rename(Year = DOL.Year,StatWk = Stat.Week,Gear = Gear.Name,Species = Species.Name, StatArea = Stat.Area,Source = Harvest.Name) %>% 
  group_by(Year,StatWk,Gear,Species,StatArea,Source)  %>% 
  summarize(CommercialHarvest = sum(Number.Of.Animals..sum.)) %>% 
  mutate(Year = as.factor(Year),StatWk = as.factor(StatWk),Gear = as.factor(Gear),StatArea = as.factor(StatArea),Species = as.factor(Species),Source = as.factor(Source))

sample.df <- sample %>% 
  filter(HarvType %in% c("CCP","HCR")) %>% 
  select(-c(ID,Agency,SampleDate,SampleType,HarvLoc,District)) %>% 
  mutate(Year = as.factor(Year),StatWk = as.factor(StatWk),Gear = as.factor(Gear),SubDist = as.factor(SubDist),Species = as.factor(Species),HarvType = as.factor(HarvType)) %>% 
  filter(!(HomerID == "18CPTpinks")) # Remove samples related to a Pink Salmon stomach contents pilot study

sampled.harvest <- full_join(harvest.df,sample.df,  # joining sample data frame to harvest data frame
              by = join_by(Year == Year,
                           StatWk == StatWk,
                           Gear == Gear,
                         StatArea == SubDist,
                         Species == Species,
                         Source == HarvType))
```


## Change Sample Area
Some samples were identified as having a mistake made in the recording of their SubDistrict (StatArea) on the ADFG sampling form. These mistakes are rectified using the ```ChangeStatArea.csv``` file to make corrections to the ```harvest``` R datatable.

```{r changestatarea}
sampled.harvest.CHANGESTATAREA <- sampled.harvest %>% 
  filter(!(HomerID %in% c("18GPG03S","18GPG06S","18GPG11S","18PSD03S","19GSD086P","20GBT063S","21GMX066S","21GMX067P") & is.na(CommercialHarvest))) %>% 
  mutate(HomerID = ifelse(Year == "2019" & StatWk == 31 & Gear == "SGN" & Species == "Pink","19GSD086P",HomerID),
         HomerID = ifelse(Year == "2018" & StatWk == 24 & Gear == "SGN" & Species == "Sockeye" & StatArea == "24130","18GPG03S",HomerID),
         HomerID = ifelse(Year == "2018" & StatWk == 25 & Gear == "SGN" & Species == "Sockeye" & StatArea == "24130","18GPG06S",HomerID),
         HomerID = ifelse(Year == "2018" & StatWk == 26 & Gear == "SGN" & Species == "Sockeye" & StatArea == "24130","18GPG11S",HomerID),
         HomerID = ifelse(Year == "2018" & StatWk == 25 & Gear == "PS" & Species == "Sockeye" & StatArea == "24190","18PSD03S",HomerID),
         HomerID = ifelse(Year == "2020" & StatWk == 31 & Gear == "SGN" & Species == "Sockeye" & StatArea == "24106","20GBT063S",HomerID)) %>% 
  mutate(StatArea = ifelse(HomerID == "18GBT02S" & is.na(CommercialHarvest),"24117",as.character(StatArea)),
         StatArea = ifelse(HomerID == "18PSD06S" & is.na(CommercialHarvest),"24190",StatArea),
         StatArea = ifelse(HomerID == "20PSD041S" & is.na(CommercialHarvest),"24191",StatArea),
         CommercialHarvest = ifelse(HomerID == "18GBT02S" & is.na(CommercialHarvest),250,CommercialHarvest),
         CommercialHarvest = ifelse(HomerID == "18PSD06S" & is.na(CommercialHarvest),5160,CommercialHarvest),
         CommercialHarvest = ifelse(HomerID == "20PSD041S" & is.na(CommercialHarvest),3693,CommercialHarvest))
```


## Missing Harvest

Other samples where a HomerID exists (strata was sampled by ADFG staff) but there is not a corresponding commercial harvest harvest record from that stratum (subdistrict, year, week, gear-type). The harvest that the sample was taken from was reported in a different subdistrict. Thus, we must add a catch record for the strata that was sampled and then subtract a corresponding catch record from a neighboring subdistrict to keep the final catch total consistent.

### Add Harvest

First determine the sample ID's when CommercialHarvest wasn't reported and determine how many samples were taken for each ID:

```{r addmissingharvest}
missing.harvest.ID <- filter(sampled.harvest.CHANGESTATAREA, is.na(CommercialHarvest))$HomerID

pink.missing.harvest <- pink.clean %>% 
  filter(HomerSampleID %in% missing.harvest.ID) %>% 
  group_by(HomerSampleID) %>% 
  summarise(MissingHarvest = Preps)

reds.missing.harvest <- reds.clean %>% 
  filter(HomerSampleID %in% missing.harvest.ID) %>% 
  group_by(HomerSampleID) %>% 
  summarise(MissingHarvest = Preps)

add.missing.harvest <- full_join(pink.missing.harvest,reds.missing.harvest)
```

### Subtract Harvest

Harvest was added to strata that were sampled but no commercial catch was reported. However, we need to subtract the same total from the harvest of a neighboring subdistrict in the same strata (year, week, gear-type). 

``` {r subtractmissingharvest}

# Manually assign which statistical area (SubDist) that the added harvest for a sample ID (HomerID) should be subtracted from based upon our knowledge of where a catch was likely to have been reported from:
sample.missing <- filter(sample, HomerID %in% missing.harvest.ID) %>% 
  mutate(SubStatArea = ifelse(HomerID == "18PSD13S",24190,NA),
         SubStatArea = ifelse(HomerID == "18PSD25S",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "18PSD26P",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "18PSD37P",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "18PSD43P",24115,SubStatArea),
         SubStatArea = ifelse(HomerID == "19PSD088S",24191,SubStatArea),
         SubStatArea = ifelse(HomerID == "19PSD089P",24191,SubStatArea),
         SubStatArea = ifelse(HomerID == "20PSD012S",24190,SubStatArea),
         SubStatArea = ifelse(HomerID == "21PSD062P",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "21PSD068S",24190,SubStatArea),
         SubStatArea = ifelse(HomerID == "21PSD069P",24190,SubStatArea),
         SubStatArea = ifelse(HomerID == "21PSD071P",24106,SubStatArea),
         SubStatArea = ifelse(HomerID == "22PSD009S",24191,SubStatArea),
         SubStatArea = ifelse(HomerID == "22PSD039S",24193,SubStatArea),
         SubStatArea = ifelse(HomerID == "22PSD040P",24193,SubStatArea))


## Create data frame that has the final commercial harvest value for the strata that are reduced to account for allocating the sample from unreported strata   
subtract.missing.harvest <- harvest %>% 
  select(-c(ID,DOL.Month,District,Subdistrict,Stat.Area.Name,Landed.Weight..sum.,Business.Name)) %>% 
  filter(Species.Name %in% c("salmon, sockeye","salmon, pink"),Gear.Name %in% c("Purse seine","Set gillnet"),Harvest.Name %in% c("State managed fishery","Private Hatchery - Fishing")) %>% 
  mutate(Gear.Name = if_else(Gear.Name == "Purse seine", "PS", Gear.Name),
         Gear.Name = if_else(Gear.Name == "Set gillnet", "SGN", Gear.Name),
         Species.Name = if_else(Species.Name == "salmon, sockeye", "Sockeye", Species.Name),
         Species.Name = if_else(Species.Name == "salmon, pink", "Pink", Species.Name),
         Harvest.Name = if_else(Harvest.Name == "State managed fishery", "CCP", Harvest.Name),
         Harvest.Name = if_else(Harvest.Name == "Private Hatchery - Fishing", "HCR", Harvest.Name)) %>% 
  filter(Gear.Name != "SGN") %>% 
  group_by(DOL.Year,Stat.Week,Gear.Name,Species.Name,Harvest.Name,Date.of.Landing,Stat.Area)  %>% 
  summarize(CommercialHarvest = sum(Number.Of.Animals..sum.)) %>% 
  right_join(sample.missing,
             by = join_by(Species.Name == Species,
                          Stat.Area == SubStatArea,
                          Date.of.Landing == SampleDate,
                          Harvest.Name == HarvType)) %>% 
  full_join(add.missing.harvest,
            by = join_by(HomerID == HomerSampleID)) %>%
  mutate(Final.CommercialHarvest = CommercialHarvest - MissingHarvest) %>%
  select(DOL.Year,Stat.Week,Gear.Name,Species.Name,Harvest.Name,Date.of.Landing,Stat.Area,Final.CommercialHarvest)

## Recreate harvest data frame. Join with the data frame above (subtract.missing.harvest) that accounts for the reallocation of harvest to unreported strata
harvest.df <- harvest %>% 
  select(-c(ID,DOL.Month,District,Subdistrict,Stat.Area.Name,Landed.Weight..sum.,Business.Name)) %>% 
  filter(Species.Name %in% c("salmon, sockeye","salmon, pink"),Gear.Name %in% c("Purse seine","Set gillnet"),Harvest.Name %in% c("State managed fishery","Private Hatchery - Fishing")) %>% 
  mutate(Gear.Name = if_else(Gear.Name == "Purse seine", "PS", Gear.Name),
         Gear.Name = if_else(Gear.Name == "Set gillnet", "SGN", Gear.Name),
         Species.Name = if_else(Species.Name == "salmon, sockeye", "Sockeye", Species.Name),
         Species.Name = if_else(Species.Name == "salmon, pink", "Pink", Species.Name),
         Harvest.Name = if_else(Harvest.Name == "State managed fishery", "CCP", Harvest.Name),
         Harvest.Name = if_else(Harvest.Name == "Private Hatchery - Fishing", "HCR", Harvest.Name)) %>% 
  group_by(DOL.Year,Stat.Week,Gear.Name,Species.Name,Harvest.Name,Date.of.Landing,Stat.Area)  %>% 
  summarize(CommercialHarvest = sum(Number.Of.Animals..sum.)) %>% 
  full_join(subtract.missing.harvest,
            by = join_by(DOL.Year, Stat.Week, Gear.Name, Species.Name, Harvest.Name, Date.of.Landing, Stat.Area)) %>% 
  mutate(CommercialHarvest = ifelse(!is.na(Final.CommercialHarvest),Final.CommercialHarvest,CommercialHarvest)) %>% 
  select(-Final.CommercialHarvest) %>% 
  rename(Year = DOL.Year, StatWk = Stat.Week, Gear = Gear.Name, Species = Species.Name, StatArea = Stat.Area, Source = Harvest.Name) %>% 
  group_by(Year,StatWk,Gear,Species,StatArea,Source) %>% 
  summarise(CommercialHarvest = sum(CommercialHarvest)) %>% 
  mutate(Year = as.factor(Year),StatWk = as.factor(StatWk),Gear = as.factor(Gear),StatArea = as.factor(StatArea),Species = as.factor(Species),Source = as.factor(Source))

## Rejoin harvest dataframe with sample dataframe
sampled.harvest.SUBTRACTED <- full_join(harvest.df,sampled.harvest.CHANGESTATAREA,  # joining sample data frame to harvest data frame
              by = join_by(Year == Year, 
                           StatWk == StatWk,
                           Gear == Gear,
                         StatArea == StatArea,
                         Species == Species,
                         Source == Source)) %>% 
  mutate(CommercialHarvest = CommercialHarvest.x) %>% 
  select(-c(CommercialHarvest.x,CommercialHarvest.y))

## Add in samples for strata that were sampled but no catch was reported:
sampled.harvest.ADDED <- sampled.harvest.SUBTRACTED %>% 
  full_join(add.missing.harvest, by = join_by(HomerID == HomerSampleID)) %>% 
  mutate(CommercialHarvest = ifelse(is.na(CommercialHarvest),MissingHarvest,CommercialHarvest)) %>% 
  select(-MissingHarvest) %>% 
  mutate(CommercialHarvest = ifelse(Year == "2021" & StatWk == "31" & Gear == "PS" & Species == "Pink" & StatArea == 24191,55+76,CommercialHarvest)) ## Individual edit of data where two HomerID samples were added to the same strata commercial harvest

```

## Sampled Harvest Data
``` {r cleansampledharvest}
sampled.harvest.CLEAN <- sampled.harvest.ADDED %>% # Rename sampled.harvest data frame to more intuitive name
  mutate(StatArea = as.factor(StatArea)) %>% 
  filter(str_detect(StatArea,"^241"))
```

# 1 - Estimate of C.hat~hi~

The otolith-derived estimate of the contribution of hatchery-area *h* to district-period stratum *i*, *C~hi~* is
$$\hat{C}_{hi} = \frac{o_{hi}}{n_i}N_i$$
where,

*h* = Hatchery-area (one of LCI, PWS, KOD, or Other)

*o~hi~* = Number of otoliths from hatchery district *h* in sample *n~i~*,

*n~i~* = Number of otoliths sampled from stratum *i* (usually 96), and

*N~i~* = Number of fish caught in stratum *i*.

The variance estimate of *C~hi~* is
$$var(\hat{C}_{hi}) = N_i^2\frac{1}{n_i-1}\frac{o_{hi}}{n_i}(1-\frac{o_{hi}}{n_i})$$

```{r C.hat_hi, message=FALSE}

# Create table of # of otoliths from hatchery-area (h) in sample from a given stratum (n_i):
sample.o_hi <- otolith.TM.data %>% 
  pivot_longer(cols = c("LCI","PWS","KOD","Other"), names_to = "HatcheryArea", values_to = "MarkedOtoliths") %>% # Pivot table to LONG format, with one count of the # of MarkedOtoliths per Strata*HatcheryArea combination
  rename(HomerID = HomerSampleID) %>% 
  mutate(HatcheryArea = as.factor(HatcheryArea)) %>% 
  group_by(Year,Gear,Species,Source,StatArea,HatcheryArea) %>% 
  summarise(o_hi = sum(MarkedOtoliths))

# Create table of total # of otoliths sampled from a given stratum (n_i):
sample.n_i <- otolith.TM.data %>% 
  group_by(Year,Gear,Species,Source,StatArea) %>% # Group into stratum (i)
  summarise(n_i = sum(NotMarked) + sum(Marked)) # SampledOtoliths = total # of otoliths sampled

# Create table of total # of fish commercially harvested in stratum (i):
harvest.N_i <- sampled.harvest.CLEAN %>%
  group_by(Year,StatWk,Gear,Species,Source,StatArea) %>% 
  summarise(N_i = sum(unique(CommercialHarvest)))

# Create a table of the otolith-derived estimated of the contribution of hatchery-area (h) to period-gear-district stratum (i):
C.hat_hi <- sample.o_hi %>% 
  full_join(sample.n_i,
            by = join_by(Year,Gear,Species,Source,StatArea)) %>% 
  full_join(harvest.N_i,
            by = join_by(Year,Gear,Species,Source,StatArea)) %>% 
  filter(!is.na(HatcheryArea)) %>% # Removing instances where HatcheryArea = NA, aka when a stratum had commercial harvest occur but no sampling occurred
  mutate(ProportionMarked = o_hi/n_i, # Calculate hatchery proportion of sample
         C.hat_hi = ProportionMarked*N_i,# EQUATION 1: Multiply hatchery proportion by the total commercial harvest for estimate hatchery proportion in commercial catch
         C.hat_hi_VAR = (N_i^2)*(1/(n_i-1))*(o_hi/n_i)*(1 - (o_hi/n_i))) # EQUATION 2: Variance estimate of the hatchery-derived commercial harvest contribution

```

```{r C.hat_hi.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_hi, desc(C.hat_hi)),10))
```

# 2 - Estimate of C.hat~Sh~

Otolith-derived estimates of the contribution of hatchery-area *h*, *C~Sh~*, to all sampled harvests (in a given StatWk but combined across all StatArea's) will be calculated as
$$\hat{C}_{Sh}=\sum_{i=1}^{Q}C_{hi}$$
where,

*Q* = Number of recovery strata in which otoliths from hatchery-area *h* are found.

The variance estimate of *C.hat~Sh~* is
$$var(\hat{C}_{Sh})=\sum_{i=1}^{Q}var(C_{hi})$$

```{r C.hat_Sh}
C.hat_Sh <- C.hat_hi %>% 
  group_by(Year,Gear,Species,Source,HatcheryArea) %>% 
  summarize(C.hat_Sh = sum(C.hat_hi), # EQUATION 3: Estimate of contribution of hatchery-area (h) to all sampled harvests within a StatWk, combined across all sampled StatArea's.
            C.hat_Sh_VAR = sum(C.hat_hi_VAR)) # EQUATION 4: Variance estimate of C.hat_Sh
```

```{r C.hat_Sh.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_Sh, desc(C.hat_Sh)),10))
```

# 3 - Estimate of C.hat~Uh~

The contribution of hatchery-area *h* to unsampled strata, *C~Uh~*, will be estimated from contribution rates associated with strata sampled from the same district-period openings as the unsampled strata using methods similar to those used for coded wire tags (Riffe et al. 1996):
$$\hat{C}_{Uh}=\sum_{i=1}^{U}\left[N_i\left(\frac{\sum_{j=1}^{S}\hat{C}_{hj}}{\sum_{j=1}^{S}N_j}\right)\right]$$
where,

*U* = Number of unsampled strata,

*N~i~* = Number of fish in *i*^th^ unsampled stratum,

*S* = Number of strata sampled in the period in which the unsampled stratum resides,

*C~hj~* = Contribution of thermal mark *h* to the sampled stratum *j*, and

*N~j~* = Number of fish in *j*^th^ sampled stratum.

The variance estimate of *C.hat~Uh~* is

$$var(\hat{C}_{Uh})=\sum_{i=1}^{U}\left[N_{i}^2\left(\frac{1}{\sum_{j=1}^{S}N_{j}-1}\right)\left(\frac{\sum_{j=1}^{S}\hat{C}_{hj}}{\sum_{j=1}^{S}N_j}\right)\left(1-\frac{\sum_{j=1}^{S}\hat{C}_{hj}}{\sum_{j=1}^{S}N_j}\right)\right]$$

```{r C.hat_Uh}
unsampled.N_i <- sample.o_hi %>%
  full_join(sample.n_i,
            by = join_by(Year,Gear,Species,Source,StatArea)) %>%
  full_join(harvest.N_i,
            by = join_by(Year,Gear,Species,Source,StatArea)) %>%
  filter(is.na(HatcheryArea)) %>% # Selecting instances where HatcheryArea = NA, aka when a stratum had commercial harvest occur but no sampling occurred
  select(-c(HatcheryArea,o_hi,n_i))

unsampled.C.hat_Uh <- C.hat_hi %>% 
  group_by(Year,Gear,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_hj = sum(C.hat_hi), N_j = sum(N_i),ProportionMarked = C.hat_hj/N_j)

C.hat_Uh <- unsampled.N_i %>% 
  left_join(unsampled.C.hat_Uh,
            by = join_by(Year,Gear,Species,Source),
            relationship = "many-to-many") %>% 
  group_by(Year,Gear,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_Uh = sum(N_i)*unique(ProportionMarked), # EQUATION 5: Estimate the contribution of hatchery-area (h) to unsampled StatArea's
            C.hat_Uh_VAR = as.integer(sum(N_i^2 * (1/(sum(N_j)-1)) * (sum(C.hat_hj)/sum(N_j)) * (1 - (sum(C.hat_hj)/sum(N_j)))   ))) %>% # EQUATION 6: Variance estimate of C.hat_Uh
  filter(!is.na(HatcheryArea))

# NA's occur in the above data frame (C.hat_Uh) when there is commercial harvest from a stratum (Year_StatWk_Gear,Species,Fishery) but no fish were sampled in that stratum (across any StatArea's) from which to estimate hatchery proportion...
# These unsampled stratum observations are removed from the C.hat_Uh data frame...

# And these unsampled stratum observations are subsetted to their own data frame: 
unsampled.stratum <- unsampled.N_i %>% 
  left_join(unsampled.C.hat_Uh,
            by = join_by(Year,Gear,Species,Source),
            relationship = "many-to-many") %>% 
  group_by(Year,Gear,Species,Source,HatcheryArea) %>% 
  summarise(C.hat_Uh = sum(N_i)*unique(ProportionMarked),
            C.hat_Uh_VAR = sum(N_i^2 * (1/(sum(N_j)-1)) * (sum(C.hat_hj)/sum(N_j)) * (1 - (sum(C.hat_hj)/sum(N_j)))   )) %>% 
  filter(is.na(HatcheryArea)) %>% 
  select(Year,Gear,Species,Source) %>% 
  unite("ID",c(Year,Gear,Species,Source), sep = "_")
# And the hatchery contribution may be estimated to these observations in Part 6 -- maybe...

```

```{r C.hat_Uh.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_Uh, desc(C.hat_Uh)),10))
```

# 4 - Estimate of C.hat~h~

An estimate of the contribution by hatchery-area *h* to all strata, sampled and unsampled, will be calculated by
$$\hat{C}_h=\hat{C}_{Sh}+\hat{C}_{Uh}$$

The variance estimate of C.hat~h~ is
$$var(\hat{C}_h)=var(\hat{C}_{Sh})+var(\hat{C}_{Uh})$$

```{r C.hat_h}
C.hat_h <- C.hat_Sh %>% 
  full_join(C.hat_Uh,
            by = join_by(Year,Gear,Species,Source,HatcheryArea)) %>% 
  replace_na(list(C.hat_Uh = 0, C.hat_Uh_VAR = 0)) %>% 
  mutate(C.hat_h = C.hat_Sh + C.hat_Uh, # EQUATION 7: Estimate of contribution of hatchery-area (h) to all strata
         C.hat_h_VAR = C.hat_Sh_VAR + C.hat_Uh_VAR) # EQUATION 8: Variance estimate of C.hat_h
```

```{r C.hat_h.table, echo=FALSE}
knitr::kable(head(arrange(C.hat_h, desc(C.hat_h)),10))
```

# 5 - Estimate of C.hat~H~

The overall hatchery contribution by all hatcheries is calculated by

$$\hat{C}_H=\sum_{h}^{H}\hat{C}_h$$
```{r C.hat_H}
C.hat_H.HATCHAREA <- C.hat_h %>% 
  group_by(Species,Gear,Year,Source,HatcheryArea) %>% 
  summarise(C.hat_H = sum(C.hat_h),
            C.hat_H_VAR = sum(C.hat_h_VAR)) %>% 
  mutate(C.hat_H_se = as.integer(sqrt(C.hat_H_VAR)),
         lower95CI = C.hat_H - 1.96*C.hat_H_se,
         upper95CI = C.hat_H + 1.96*C.hat_H_se)

C.hat_H.TOTAL <- C.hat_h %>% 
  group_by(Species,Gear,Year,Source) %>% 
  summarise(C.hat_H = sum(C.hat_h),
            C.hat_H_VAR = sum(C.hat_h_VAR)) %>% 
  mutate(C.hat_H_se = as.integer(sqrt(C.hat_H_VAR)),
         lower95CI = C.hat_H - 1.96*C.hat_H_se,
         upper95CI = C.hat_H + 1.96*C.hat_H_se)
```

```{r C.hat_H.table, echo=FALSE}
knitr::kable(C.hat_H.TOTAL)
```

# Catch Proportion Comparison

```{r catch.comparison}
catch.comparison.HATCHAREA <- harvest.N_i %>% 
  unite("ID",c(Year,Gear,Species,Source), sep = "_") %>% 
  filter(!(ID %in% unsampled.stratum$ID)) %>% 
  separate(ID, into = c("Year","Gear","Species","Source"), sep = "_") %>% 
  group_by(Species,Gear,Year,Source) %>% 
  summarize(Catch = sum(N_i)) %>% 
  left_join(C.hat_H.HATCHAREA,
            by = join_by(Species,Gear,Year,Source)) %>% 
  mutate(HatcheryProportion = C.hat_H/Catch)

catch.comparison.TOTAL <- harvest.N_i %>% 
  unite("ID",c(Year,Gear,Species,Source), sep = "_") %>% 
  filter(!(ID %in% unsampled.stratum$ID)) %>% 
  separate(ID, into = c("Year","Gear","Species","Source"), sep = "_") %>% 
  group_by(Species,Gear,Year,Source) %>% 
  summarize(Catch = sum(N_i)) %>% 
  left_join(C.hat_H.TOTAL,
            by = join_by(Species,Gear,Year,Source)) %>% 
  mutate(HatcheryProportion = C.hat_H/Catch)
```

```{r catchcomp.table, echo=FALSE}
knitr::kable(catch.comparison.TOTAL)
```

